---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(knitr)
library(kableExtra)
library(tidyverse)
```

# DoxyICU Results
This repo provides the raw data and analysis code to support the paper <em>Doxycycline for the prevention of progression of COVID-19 to severe disease requiring intensive care unit (ICU) admission: a randomized, controlled, open-label, parallel group trial (DOXPREVENT.ICU)</em> by Dhar et al, submitted to the Lancet on 26 December 2021.

The raw data are contained in `data 2021-12 Dec-24.Rds`, the analysis code in `analysis.Rmd` and the results in `analysis.html`.

# Data description
```{r, echo=FALSE}
data <- readRDS("data 2021-12 Dec-24.Rds")

types <- data %>% 
           dplyr::summarise_all(class) %>% 
           tidyr::gather(variable, class)

labels <- tribble(~variable, ~Label,
"UUID", "Unique database ID",
"PatNo", "Patient number. 4 characters left-padded with zeroes.  Numbered from 1 within site",
"SiteID", "Site ID. 3 characters left-padded with zeroes.",
"Treatment", "Randomised treatment: `Standard of Care`, `Doxycycline`",
"ICDate", "Date of informed consent",         
"Sex", "Patient sex: 'Male'=1, 'Female'=2 or NA",
"Age", "Patient age at baseline in completed years",                      
"Ethnicity", "Patient ethnicity: 'White'=1, 'Black Caribbean'=2, 'Black African'=3  'Indian'=4, 'Pakistani'=5, 'Bangladeshi'=6, 'Chinese'=7, 'Other'=8",
"Temperature", "Patient temperature at baseline in Celsius",                    
"TemperatureLocation", "Location at which temperature was measured: 'On forehead'=1, 'Orally'=2",
"OnsetDate", "Date of symptom onset",                      
"AdmissionDate", "Date of hospital admission",
"FinalStatus", "Final status: 'Completed 14 days of treatment and discharged'=1, 'Discharged before completing 14 days of treatment'=2, 'Discontinued treatment because of treatment-related adverse events'=3, 'Discontinued treatment for reasons unrelated to treatment'=4, 'Lost to follow-up'=5, 'Consent withdrawn'=6, 'Other'=7, 'Died'=8",                    
"ICUAdmissionNeeded", "Did the patient need admission to ICU? 'Yes'=1, 'No'=0",
"ICUAdmissionSuccess", "Was the patient actually admitted to ICU? 'Yes'=1, 'No'=0",            
"ICUAdmissionNeededDate", "Date on which the patient's need for ICU admission was determined",
"ICUNonAdmissionReason", "Why was the patient NOT admitted to ICU? 'Lack of beds'=1, 'Patient unsuitable'=2, 'Patient declined'=3, 'Other'=4",          
"ICUDischargeDate", "Date of ICU discharge",
"HospitalDischargeDate", "Date of hospital discharge",          
"Alive", "Was the patient alive on the date of last follow-up? 'Yes'=1, 'No'=0",
"DateOfDeath", "Date of death",                    
"RelatedDeath", "Was the death related to COVID-19? 'Yes'=1, 'No'=0",
"FollowUpDate", "Date of last follow-up",                   
"PCRTest", "Inclusion/exclusion: Positive PCR test for SARS-CoV-2?  'Yes'=1, 'No'=0",
"Hypersensitivity", "Inclusion/exclusion: Known hypersensitivity to doxycycline? 'Yes'=1, 'No'=0",               
"MyastheniaGravis", "Inclusion/exclusion: Myasthenia gravis?  'Yes'=1, 'No'=0",
"Pregnancy", "Inclusion/exclusion: Pregnancy?  'Yes'=1, 'No'=0 'Not applocable'=2",                      
"Cirrhosis", "Inclusion/exclusion: Cirrhosis?  'Yes'=1, 'No'=0",
"Dyspnea", "Inclusion/exclusion: New continuous cough or dyspnea?  'Yes'=1, 'No'=0",                        
"Diarrhoea", "Inclusion/exclusion: Diarrhoea?  'Yes'=1, 'No'=0",
"Anosmia", "Inclusion/exclusion: Has the patient recently lost their sense of smell? 'Yes'=1, 'No'=0",                        
"Pneumonia", "Inclusion/exclusion: Evidence of pneumonia on x-ray?  'Yes'=1, 'No'=0 'X-ray not done`=2",
"Rash", "Inclusion/exclusion: Does the patient have 'COVID rash' or 'COVID fingers and toes'? 'Yes'=1, 'No'=0",                           
"TrtDoxycycline", "Post-baseline treatment with doxycycline? 'Yes'=1, 'No'=0",
"TrtAntibiotics", "Post-baseline treatment with other antibiotics? 'Yes'=1, 'No'=0",                 
"TrtAnalgesics", "Post-baseline treatment with analgesics? 'Yes'=1, 'No'=0",
"TrtDiabetes", "Post-baseline treatment for diabetes? 'Yes'=1, 'No'=0",                    
"TrtHeartFailure", "Post-baseline treatment for heart failure? 'Yes'=1, 'No'=0",
"TrtAntiHypertensives", "Post-baseline treatment with anti-hypertensives? 'Yes'=1, 'No'=0",           
"TrtAntiDepressants", "Post-baseline treatment with anti-depressants? 'Yes'=1, 'No'=0",
"TrtCancer", "Post-baseline treatment for cancer? 'Yes'=1, 'No'=0",                      
"TrtImmunosuppressant", "Post-baseline treatment with immunosuppressants? 'Yes'=1, 'No'=0",
"TrtSteroids", "Post-baseline treatment with steroids? 'Yes'=1, 'No'=0",                    
"TrtAntiCoagulantsProphylactic", "Post-baseline prophylactic treatment with anti-coagulants? 'Yes'=1, 'No'=0",
"TrtAntiCoagulantsTherapeutic", "Post-baseline therapeutic treatment with anti-coagulants? 'Yes'=1, 'No'=0",   
"TrtFavipiravir", "Post-baseline treatment with favipiravir? 'Yes'=1, 'No'=0",
"TrtHydroxychloroquine", "Post-baseline treatment with hydroxychloroquine? 'Yes'=1, 'No'=0",          
"TrtRemdesivir", "Post-baseline treatment with remdesivir? 'Yes'=1, 'No'=0",
"TrtTocilizumab", "Post-baseline treatment with tocilizumab? 'Yes'=1, 'No'=0",                 
"TrtIvermectin", "Post-baseline treatment with vermectin? 'Yes'=1, 'No'=0",
"TrtOther", "Other relevant post-baseline treatment?Discontinuation due to ",                       
"TreatmentDuration", "Treatment duration: 14 values from '1 day' to '14 days', '>14 days'",
"TreatmentDiscontinuationReason", "Reason for dsiscontinuation of treatment: 'Patient improved, was discharged or the patient/physician chose not to give the full 14 day course of treatment'='1', 'Patient experienced unacceptable treatment-related side effects'='2', 'Other reasons'='3'", 
"AENausea", "Discontinuation due to nausea? 'Yes'=1, 'No'=0",
"AECramps", "Discontinuation due to cramps? 'Yes'=1, 'No'=0",                       
"AEDiarrhoea", "Discontinuation due to diarrhoea? 'Yes'=1, 'No'=0",
"AETinnitus", "Discontinuation due to tinnitus? 'Yes'=1, 'No'=0",                     
"AEGenitalSoreness", "Discontinuation due to genital soreness? 'Yes'=1, 'No'=0",
"AERash", "Discontinuation due to rash? 'Yes'=1, 'No'=0",                         
"AEHeadache", "Discontinuation due to headache? 'Yes'=1, 'No'=0",
"AEVision", "Discontinuation due to vision effects? 'Yes'=1, 'No'=0",                       
"AEPalpitation", "Discontinuation due to palpitations? 'Yes'=1, 'No'=0",
"AEOther", "Discontinuation due to other reasons? 'Yes'=1, 'No'=0",                        
"StratumILD", "Risk factor for stratification: ILD 'Yes'=1, 'No'=0",
"StratumCOPD", "Risk factor for stratification: COPD 'Yes'=1, 'No'=0",                    
"StratumBronchiectasis", "Risk factor for stratification: bronchiectasis 'Yes'=1, 'No'=0",
"StratumAsthma", "Risk factor for stratification: asthma 'Yes'=1, 'No'=0",                  
"StratumOtherLung", "Risk factor for stratification: other lung condition 'Yes'=1, 'No'=0",
"StratumDiabetes", "Risk factor for stratification: diabetes 'Yes'=1, 'No'=0",                
"StratumHeartDisease", "Risk factor for stratification: heart disease 'Yes'=1, 'No'=0",
"StratumHypertension", "Risk factor for stratification: hypertension 'Yes'=1, 'No'=0",            
"StratumCancer", "Risk factor for stratification: cancer 'Yes'=1, 'No'=0",
"StratumOther", "Risk factor for stratification: other 'Yes'=1, 'No'=0",                   
"StratumCode", "Derived stratum code: 1, 2 or 3",
"Stratum", "Derived stratum text: NONE [0], LUNG [1], OTHER [2]")

types %>% 
  left_join(labels, by="variable") %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
  
```